---
title: "myope2"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Code and simulated data for the article **Age and sex-specific risks of myocarditis and pericarditis following Covid-19 messenger RNA Vaccines** by Le Vu et al., available at https://www.nature.com/articles/s41467-022-31401-5

```{r load}
source("functions.R")
```

# Population
- We simulate a population of size `npop` with a base prevalence of the outcome `p0` and exposed to two vaccines, V1 and V2, according to the ranking of the first two doses, D1 or D2.
- A vector `pe` of four probabilities defines exposures levels to the combination of vaccines and doses. 
- Note that in our study, exposure refers to being within a window of 21 days from the last dose of vaccine when included in the case-control sample.
- Relative risks are parameterized like so :
    - `risk_v1` risk of outcome associated with exposure to V1 relative to being unexposed.
    - `risk_v2_v1` risk of outcome associated with exposure to V2 relative to v1.
    - `risk_d2` risk of outcome associated with exposure to the second dose relative to the first, applied for both vaccines.

```{r popsim}
set.seed(1234)
n <-  32e6
pe <- setNames(c(0.08, 0.08, 0.01, 0.01), c("V1_D1", "V1_D2", "V2_D1", "V2_D2"))
pop0 <- make_population(npop = n, 
                        p0 = 1e-4,
                        pe = pe,
                        risk_v1 = 1.5,
                        risk_d2 = 2,
                        risk_v2_v1 = 2)

```

- Description of the population
```{r despop}
with(pop0, table(expo, case))

```

# Case-control study
- We now sample `ncase` cases and `ra * ncase` (matched) controls.

```{r ccsim}
df_cc <- make_cc_study(pop0, ncase = 1500, ra = 10)

```

- Description of the study sample
```{r descc}
(t0 <- with(df_cc, table(expo, case)) )

```

## Measure of association
- We estimate the odds ratios by conditional logistic regression
```{r}
require(survival)
mod <- clogit(case ~ expo + strata(set), data = df_cc)
r0 <- cbind( OR = round(exp(coef(mod)), 3),
               round(exp(confint(mod)), 3) )
rownames(r0) <- gsub("expo", "", rownames(r0))
res <- merge(unclass(t0), r0, by = 0, all.x = TRUE)
colnames(res) <- c("expo", "ctl", "case", "or", "lo", "up")
res
```

## Vaccine-associated cases
- Given a vector `pd` of proportion of doses received, we then estimate
    - the population attributable fraction `paf`,
    - the excess of vaccine-associated cases `ec` per `nper` doses, and
    - the number of doses needed for the occurrence of one vaccine-associated case `nnh`.
- Here, we assume that the associations are causal, that the relative risks can reasonably be estimated by the odds ratios,  and that each vaccinee has experienced the (short) window of exposure, so that exposure is given by the number of respective doses received by the population.
- We derive the confidence intervals according to Greenland, 1987 [1].

```{r}
pd <- setNames( c(0.65, 0.59, 0.09, 0.08), names(pe) )
res <- cbind(res, nd = c(NA, pd * n) )
nper <- 1e5 # denominator
```


```{r}
u_case <- res[1, "case"] ## unexposed cases
ar_ic0 <- with(res[-1,], ar_delta_greenland(RR = or, RR_lo = lo,
                                              RR_up = up, n1 = case,
                                              n0 = u_case, type = 2))

paf_ic0 <- matrix(ar_ic0$ci_logitbased_greenland, ncol = 3, byrow = FALSE)
ec_ic0 <- matrix(
    with(res[-1,], ar_ic0$ci_logitbased_greenland * (case + u_case) / nd * nper),
    ncol = 3, byrow = FALSE)
nnh_ic0 <- (nper / ec_ic0)
nnh_ic0 <- nnh_ic0[, c(1,3,2)] ## reorder
a <- matrix(cbind(paf_ic0, ec_ic0, nnh_ic0), ncol = 9, byrow = FALSE,
               dimnames = list(res[-1, 1], c("paf", "paf_lo", "paf_up",
                                       "ec", "ec_lo", "ec_up",
                                       "nnh", "nnh_lo", "nnh_up")) )
signif(a, 4)
```

# Reference
1. Greenland S. Variance estimators for attributable fraction estimates consistent in both large strata and sparse data. Statistics in Medicine. 1987;6(6):701–8. 
